---
title: "11-load-rmi"
output:
  html_notebook:
    toc: yes
    toc_depth: 3
    toc_float: yes
    number_sections: true
---

## Load libs, common functions
```{r Load rmi}
source(knitr::purl("10-load-data.Rmd"))
fs::file_delete("10-load-data.R")
pacman::p_load(DataExplorer)
```

# LDV National Stock Forecast
This contains Light Duty Vehicle (LDV) stock projections for the United States through 2050 by vehicle type. It does not contain state level stock. It will need to be grouped into RMI vehicle categories and formatted correctly.

```{r load ldv_stock_natl_frcst}

ldv_stock_natl_frcst <- read_excel("LDV_Stock_National.xlsx", skip=5)
ldv_stock_natl_frcst

```

To clean the data, we'll be dropping full NA columns and rows.  Let's check to see the effect of this:

```{r check ldv_stock_natl_frcst NAs, fig.height=8, purl=FALSE}
plot_missing(ldv_stock_natl_frcst)

```
Looks like there's no NA columns or rows. However, upon examination the 'Growth (2019-2050)' column is a character, not a dbl, and features NA values. Since this column is not needed, it will be removed, as well as other redundant columns that do not usefully identify the vehicle stock.

```{r clean ldv_stock_natl_frcst}
ldv_stock_natl_frcst <- ldv_stock_natl_frcst %>%
  clean_names() %>%
  select(-growth_2019_2050, -vehicle_category, -fuel_type, -api_key, -units)

ldv_stock_natl_frcst[] <- lapply(ldv_stock_natl_frcst, gsub, pattern='Light-Duty Vehicle Stock: ', replacement='')

ldv_stock_natl_frcst
```


# LDV Fuel Economy Forecast
This contains Light Duty Vehicle (LDV) fuel economy projections for the United States through 2050 by vehicle type. It does not consider information at the state level. It will need to be grouped into RMI vehicle categories and formatted correctly.

```{r load ldv_fuel_econ}

ldv_fuel_econ <- read_excel("LDV_Fuel_Economy.xlsx", skip=5)
ldv_fuel_econ

```

To clean the data, we'll be dropping full NA columns and rows.  Let's check to see the effect of this:

```{r check ldv_fuel_econ, fig.height=8, purl=FALSE}
plot_missing(ldv_fuel_econ)

```
Looks like there's no NA columns or rows. There are no redundant columns nor is any data formatted incorrectly. The vehicle type column is renamed to match the last df.


```{r clean ldv_fuel_econ}
ldv_fuel_econ <- ldv_fuel_econ %>%
  clean_names() %>% rename(full_name = x1)

ldv_fuel_econ
```

# Suggested_VEH_Crosswalk

# LDV Fuel Economy Forecast
This contains Light Duty Vehicle (LDV) fuel economy projections for the United States through 2050 by vehicle type. It does not consider information at the state level. It will need to be grouped into RMI vehicle categories and formatted correctly.

```{r load ldv_cross}

ldv_cross <- read_excel("Suggested_VEH_Crosswalk.xlsx") 
ldv_cross <- ldv_cross %>% clean_names() %>% select(-
table_39_vehicle_category, -table_39_fuel) %>% rename(full_name = table_40)

ldv_cross

```

# Standard Vehicle Types

In this section, we standardize vehicle types across all datasets, removing redundant summary rows and adding a new column for RMI classification


```{r standard_veh_types}

ldv_cross 
ldv_fuel_econ
ldv_stock_natl_frcst 

ldv_cross_std = ldv_cross

ldv_stock_std = ldv_stock_natl_frcst %>% filter(!row_number() %in% c(3,18,19,22,37,38)) #3,18,19,22,37,38 
ldv_stock_std

ldv_fuel_std = ldv_fuel_econ %>% filter(!row_number() %in% c(17,34:42 )) #17,34:42 
ldv_fuel_std

unique(ldv_cross_std$full_name) == unique(ldv_fuel_std $full_name) #fuel and cross are same, need fix stock

# ldv_stock_std[] <- lapply(ldv_stock_std, gsub, pattern=': High oil and gas supply', replacement='')
# ldv_cross_std[] <- lapply(ldv_cross_std, gsub, pattern=' mpg gas equiv', replacement='')
# ldv_fuel_std[] <- lapply(ldv_fuel_std, gsub, pattern=' mpg gas equiv', replacement='')

ldv_stock_std[] <- lapply(ldv_stock_std, gsub, pattern=': High oil and gas supply', replacement='_STOCK')
ldv_cross_std[] <- lapply(ldv_cross_std, gsub, pattern=' mpg gas equiv', replacement='')
ldv_fuel_std[] <- lapply(ldv_fuel_std, gsub, pattern=' mpg gas equiv', replacement='_MPG')

# unique(ldv_cross_std$full_name) == unique(ldv_fuel_std $full_name) #all true
# unique(ldv_cross_std$full_name) == unique(ldv_stock_std $full_name) #all true
# unique(ldv_fuel_std$full_name) == unique(ldv_stock_std $full_name) #all true

ldv_stock_std
ldv_cross_std
ldv_fuel_std



```


# Time Series Conversion

In this section, we will eliminate data after the year 2030 and convert to time series


```{r time series, purl=FALSE}

ldv_cross_std
ldv_cross_ts = as.data.frame( t(ldv_cross_std)) %>% row_to_names(row_number = 1) %>% clean_names()
ldv_cross_ts

ldv_fuel_ts = as.data.frame( t(ldv_fuel_std)) %>% row_to_names(row_number = 1) %>% clean_names() %>% filter(!row_number() > 12)
ldv_fuel_ts

ldv_stock_ts = as.data.frame( t(ldv_stock_std)) %>% row_to_names(row_number = 1) %>% clean_names() %>% filter(!row_number() > 12) 
ldv_stock_ts

ldv_stock_ts = rownames_to_column(ldv_stock_ts, "year")
ldv_stock_ts[] <- lapply(ldv_stock_ts, gsub, pattern='x', replacement='1/1/')
#ldv_stock_ts[,1]=as.numeric(ldv_stock_ts[,1])

ldv_fuel_ts = rownames_to_column(ldv_fuel_ts, "year")
ldv_fuel_ts[] <- lapply(ldv_fuel_ts, gsub, pattern='x', replacement='1/1/')
#ldv_fuel_ts[,1]=as.numeric(ldv_fuel_ts[,1])

ldv_stock_ts =  ldv_stock_ts %>% mutate_at(-1, as.numeric)
ldv_fuel_ts = ldv_fuel_ts %>% mutate_at(-1, as.numeric)

ldv_stock_ts
ldv_fuel_ts

stock_ts = xts(ldv_stock_ts[,-1], as.Date(ldv_stock_ts$year, format='%m/%d/%Y'))
stock_ts = stock_ts*1000000 #converts from million vehicle to single vehicle units
stock_ts

fuel_ts = xts(ldv_fuel_ts[,-1], as.Date(ldv_fuel_ts$year, format='%m/%d/%Y'))
fuel_ts

class(stock_ts)
class(fuel_ts)

plot.xts(fuel_ts)
plot.xts(stock_ts)

stock_ts
fuel_ts




```



# National VMT

# 
This contains Light Duty Vehicle (LDV) and other transport mode VMT projections for the United States through 2050. It does not consider information at the state level or by vehicle type. It will need to be formatted to only store LDV data.

```{r load clean conversion}

ldv_vmt <- read_excel("Total_VMT.xlsx", skip=3, n_max=1)


ldv_vmt

#Light-Duty Vehicles less than 8,501 pounds
#unit: Billion vehicle miles traveled

ldv_vmt <- ldv_vmt %>% select(-1,-2) %>% clean_names() %>% select(1:12)
ldv_vmt

ldv_vmt_ts = as.data.frame( t(ldv_vmt)) # %>% filter(!row_number() > 12)
ldv_vmt_ts

ldv_vmt_ts = rownames_to_column(ldv_vmt_ts, "year")
ldv_vmt_ts[] <- lapply(ldv_vmt_ts, gsub, pattern='x', replacement='1/1/')

ldv_vmt_ts = ldv_vmt_ts %>% mutate_at(-1, as.numeric)

vmt_ts = xts(ldv_vmt_ts[,-1], as.Date(ldv_vmt_ts$year, format='%m/%d/%Y'))


vmt_ts = vmt_ts * 1000000000 #converts from billion vmt to vmt
vmt_ts

class(vmt_ts)

plot.xts(vmt_ts)


```



##Write to RDS
``` {r Write to RDS}
saveRDS(vmt_ts, "clean_data/vmt_ts")
saveRDS(stock_ts, "clean_data/stock_ts")
saveRDS(fuel_ts, "clean_data/fuel_ts")
saveRDS(ldv_cross_ts, "clean_data/ldv_cross_ts")
saveRDS(ldv_cross_std, "clean_data/ldv_cross_std")
```



