---
title: "32-fuel-forecast"
output: html_notebook
---

The purpose of this notebook is forecasting - using both  current features and by generating new features based upon the originals with assumptions.  The forecasts should also be tested here to ensure that they  meet expectations and assumptions about them. 

# Load cleaned data

```{r Load}
source(knitr::purl("10-load-data.Rmd"))
fs::file_delete("10-load-data.R")

#Case Study data
fuel_ts = readRDS("clean_data/fuel_ts")
vmt_ts = readRDS("clean_data/vmt_ts")

#External data
state_vmt = readRDS("clean_data/state_vmt")

#Generated data

if (GOAL == FALSE){

az_stock_adj_ts = readRDS(paste("states/",STATE_KEY,"/ldv_stock_2019_2030",sep=""))

} else {
  
az_stock_adj_ts = readRDS(paste("states/",STATE_KEY,"/ldv_stock_2019_2030_EVGOAL_",EV_TARGET,sep=""))

}
#checks to see if stock modified to meet target

#az_stock_adj_ts = readRDS(paste("states/",STATE_KEY,"/ldv_stock_2019_2030",sep=""))

fuel_ts
vmt_ts
state_vmt
az_stock_adj_ts

#Arizona is 0.02254006% of US VMT

#can calculate miles az
#can calculate miles total LDVs AZ
#CAN calc miles per LDV
#can calculate emissions
#can compare to IEA emisssions


####
#quick note

#LDVs are responsible for 0.8964 of US VMT on average
#https://www.bts.gov/browse-statistical-products-and-data/freight-facts-and-figures/vehicle-miles-traveled-highway
#2924053/3261772


####
#quick note support
#LDVs responsible for 0.8846 of US VMT on average
#https://www.fhwa.dot.gov/policyinformation/statistics/2020/vm1.cfm
#2568745/2903622

#(0.8846+0.8964)/2
#0.8905

```

# Calculate Ratio AZ VMT to US VMT

```{r ratio_vmt}

us_vmt = state_vmt %>% filter(state=="Grand Total")
value_us_vmt = sum(us_vmt$vmt)
value_us_vmt

az_vmt = state_vmt %>% filter(state=="Arizona")
value_az_vmt = sum(az_vmt$vmt)
value_az_vmt

az_vmt_ratio = value_az_vmt/value_us_vmt
az_vmt_percent = az_vmt_ratio *100
az_vmt_percent
#2.25%

```

According to data from FHWA, at `r value_az_vmt` miles, Arizona represented `az_vmt_percent`% of the total US VMT in 2020. If we make the following assumptions, we can assign vehicle miles traveled to each individual LDV by type.

## Arizona LDV VMT represents the same proportion of the state total VMT as national LDV VMT does to national total VMT

### Data from both FHWA and BTS support that national LDV VMT makes up 89.05% of all national motor vehicle VMT
### Assume same is true for Arizona
### https://www.fhwa.dot.gov/policyinformation/statistics/2020/vm1.cfm
### https://www.bts.gov/browse-statistical-products-and-data/freight-facts-and-figures/vehicle-miles-traveled-highway

###

## Each LDV travels approximately the same number of miles, regardless of type

### Supported by data from DOE that shows cars, light trucks, busses, and delivery trucks have about the same annual VMT
### https://afdc.energy.gov/data/10309

# Estimate Arizona VMT forecast

```{r vmt_forecast_az}

az_vmt_ts = vmt_ts * az_vmt_ratio
plot(az_vmt_ts)

az_vmt_ts

value_az_vmt

```


# Assign VMT to Stock

The following section assigns VMT to each Arizona vehicle by type. For each year, the average VMT per LDV is calculated. Then, each LDV type has its stock multiplied by the average VMT. The result is a time series showing total VMT per type per year.


```{r assign_vmt}

#az_stock_adj_ts = readRDS("answers/az_ldv_stock_2019_2030")

az_vmt_per_type_ts = az_stock_adj_ts

for (x in 1:12){

  ann_veh_sum = sum(az_vmt_per_type_ts[x,1:32])

  ann_vmt = az_vmt_ts[x]

  vmt_per_veh = as.numeric(ann_vmt / ann_veh_sum) #checks out with national average VMT per vehicle

  #by dividing each year's AZ VMT (ann_vmt) by the total AZ stock of vehicles that year (ann_veh_sum), we get an average vmt per LDV vehicle in AZ
  
  for (y in 1:32){
    
    az_vmt_per_type_ts[x,y] = az_vmt_per_type_ts[x,y] * vmt_per_veh
    
    #each year, we multiply each LDV stock by vmt_per_veh, yielding total vmt per LDV type per year 
  }
}
  
az_vmt_per_type_ts

value_az_ev


```

# Assign gallons per type per year using fuel economy

```{r gallons}

#fuel_ts
#az_vmt_per_type_ts

dim(fuel_ts) == dim(az_vmt_per_type_ts)
#TRUE

colnames(fuel_ts) == colnames(az_vmt_per_type_ts)
#FALSE, doesn't affect math but be aware

colnames(fuel_ts)
colnames(az_vmt_per_type_ts)

gallons_ts = az_vmt_per_type_ts/fuel_ts #divide miles by mpg to yield gallons

gallons_ts[!is.finite(gallons_ts)] <- 0 #handles NaNs

#first two years of 100 mile EV trucks throwing NaNs, follow up on this

gallons_ts
sum(gallons_ts)



```

# Assign BTU per type, based on fuel type

```{r}

#BTU per gasoline gallon :
btu_gasgl = 120476

#note, all fuel economy values are in MPGE (miles per gal gas equivalent) so we will only use this btu/gas value to calculate btu


btu_ts = gallons_ts*btu_gasgl
btu_ts

colnames(btu_ts) = gsub("_stock", "_btu", colnames(btu_ts))


```






##Write to RDS

``` {r export answer}

#saveRDS(btu_ts, "answers/btu_ts")
#saveRDS(btu_ts, paste("states/",STATE_KEY,"/ldv_btu_2019_2030",sep=""))


if (GOAL == FALSE){

saveRDS(btu_ts, paste("states/",STATE_KEY,"/ldv_btu_2019_2030",sep=""))

} else {
  
saveRDS(btu_ts, paste("states/",STATE_KEY,"/ldv_btu_2019_2030_EVGOAL_",EV_TARGET,sep=""))

}

```










































































































