---
title: "50-rmi-answers"
output: html_notebook
---

The purpose of this notebook is to load and view the time series objects for submission to the RMI US Program Technical Assessment.

All submitted graphics are generated in notebook 40-data-visualization

# Clean and load all data sets

```{r required libraries for feat eng}
source(knitr::purl("10-load-data.Rmd", quiet=TRUE))
fs::file_delete("10-load-data.R")

```

# 1A: LDV Stock Times Series

```{r}
az_stock_ts  = readRDS(paste("states/","Arizona","/ldv_stock_2019_2030",sep=""))
az_stock_ts
plot(az_stock_ts)
```



# 1B: LDV BTU Times Series

```{r}
az_btu_ts = readRDS(paste("states/","Arizona","/ldv_btu_2019_2030",sep=""))
az_btu_ts
plot(az_btu_ts)
```

# 1C: LDV CO2 Times Series

```{r}
az_co2_ts = readRDS(paste("states/","Arizona","/ldv_co2_2019_2030",sep=""))
az_co2_ts
plot(az_co2_ts)
```

# 4A: LDV Stock, EV Target, Times Series

```{r}
goal_stock_ts = readRDS(paste("states/Arizona/ldv_stock_2019_2030_EVGOAL_0.45",sep=""))
goal_stock_ts
plot(goal_stock_ts)
```

# 4B: LDV Emissions, EV Target, Time Series

```{r}
goal_co2_ts = readRDS(paste("states//Arizona//ldv_co2_2019_2030_EVGOAL_0.45",sep=""))
goal_co2_ts
plot(goal_co2_ts)
```


